A Large and Variable Leading Tail of Helium in a Hot Saturn Undergoing Runaway Inflation

### 2023-08-10 paper/paper1/ms.tex
Ohmic dissipation-dominated implies RLO, we should be more upfront about that

\subsection{Ohmic dissipation-driven mass loss}
The Ohmic dissipation mechanism made two key observable predictions.  Anomalous heating efficiency, $\epsilon$, should initially increase as a function of equilibrium temperature, then degrade as irradiation exceeds equilibrium temperatures of about 2000 K.  Second, hot Saturns ($\lesssim0.5 M_\mathrm{Jup}$) should undergo runaway evaporation, whereas hot Jupiters ($\gtrsim 1 M_\mathrm{Jup}$) should reach stable equilibrium---albeit inflated---radii after Gyr timescales.  Both of these outcomes predated data that could validate them, and these outcomes differ from the behavior of other heating mechanisms, such as photoevaporation or tides.

@@

\subsection{Weighing the causes for a lack of sub-Saturns}\label{secLackofSaturns}

When applied to an ensemble of planet systems, both the XUV irradiation and Ohmic dissipation mechanisms expect a void in the planet mass-radius plane. However, the theories make different quantitative predictions for the placement of the dividing lines between stable and unstable populations.  We illustrate these differences in Figure \ref{fig:tf2018_theory}, which shades the mass-radius plane with predictions for the inflation timescale under HAT-P-67b-like conditions.  The XUV irradiation timescale better matches the distribution of planets, with the 0.1g~cm$^{-3}$ density contour setting a conspicuous dividing line in density.  The Ohmic dissipation shading predicts short inflation timescales extending into a region of Jupiter-mass planets with radii commonly observed.  The better match of observed exoplanet demographics to the XUV irradiation shading disfavors Ohmic dissipation.

@@

Finally, we consider the prospect that the planetary mass of HAT-P67b is much lower than the 1-$\sigma$ estimate, such that the white-light planetary radius \emph{is} the Hill radius (we have previously assumed the Hill radius is 2.7 planetary radii).  This \edit1{true Roche Lobe Overflow (RLO)} scenario could manifest enormous mass loss rates, with the atmosphere's reservoir of mass fleeing the gravitational potential well directly, without the cushion of an exosphere.  \edit1{We propose that this terminal RLO should have a subtle-but-distinctive observational consequence.  Since heavy elements have smaller scale heights than hydrogen/helium, the observed ratio of heavy element to lighter element lines should be inversely correlated with the number of scale heights separating the white-light radius from the Hill radius.  All else equal, we would expect an anomalously large ratio of heavy metals compared to hydogen/Helium in the RLO case.  The extremely large Helium excess absorption, combined with the possibility of enhanced metal line depths in HAT-P-67~b---however uncertain---motivates a search for metal-line spectral tracers that may have previously been dismissed as too unlikely to detect.}

>>>>

\subsection{Ohmic dissipation-driven mass loss}\label{secOhmicD}
\edit1{Ohmic dissipation can be visualized by considering the qualitative extreme of zero XUV flux.  Here, we would predict zero radius inflation and therefore no mass loss.  The docile upper atmosphere would lose atoms one-by-one through Jeans escape, an insignificant trickle.  Instead, Ohmic dissipation heating increases the radius of the planet while preserving that low mass loss rate for a long time, until a critical moment: when the planet radius reaches the Hill radius, resulting in Roche Lobe Overflow.  So XUV irradiation and Ohmic dissipation could differ in 1) the timing and duration of their mass loss phase, and 2) the mass/radius/semi-major-axis conditions needed to expect the mass loss phase for a given system.  Since we only have a mass upper limit, HAT-P-67~b could have a mass as low as $M<0.05M_\mathrm{Jup}$, which would shrink its Hill radius to the observed planetary radius, causing the profound mass loss we see.  The prospect of this scenario motivates quantitative consideration of Ohmic dissipation.}

The Ohmic dissipation mechanism made two key observable predictions.  Anomalous heating efficiency, $\epsilon$, should initially increase as a function of equilibrium temperature, then degrade as irradiation exceeds equilibrium temperatures of about 2000 K.  Second, hot Saturns ($\lesssim0.5 M_\mathrm{Jup}$) should undergo runaway evaporation, whereas hot Jupiters ($\gtrsim 1 M_\mathrm{Jup}$) should reach stable equilibrium---albeit inflated---radii after Gyr timescales.  Both of these outcomes predated data that could validate them, and these outcomes differ from the behavior of other heating mechanisms, such as photoevaporation or tides.

@@

\subsection{Weighing the causes for a lack of sub-Saturns}\label{secLackofSaturns}
When applied to an ensemble of planet systems, both the XUV irradiation and Ohmic dissipation mechanisms expect a void in the planet mass-radius plane.  However, the theories make different quantitative predictions for the placement of the dividing lines between stable and unstable populations.  We illustrate these differences in Figure \ref{fig:tf2018_theory}, which shades the mass-radius plane with predictions for the inflation timescale under HAT-P-67b-like conditions.  The XUV irradiation timescale better matches the distribution of planets, with the 0.1g~cm$^{-3}$ density contour setting a conspicuous dividing line in density.  The Ohmic dissipation shading predicts short inflation timescales extending into a region of Jupiter-mass planets with radii commonly observed.  The better match of observed exoplanet demographics to the XUV irradiation shading disfavors Ohmic dissipation.

@@

Finally, we consider the prospect alluded to in \S \ref{secOhmicD} that the planetary mass of HAT-P67~b is much lower than the 1-$\sigma$ estimate, such that the white-light planetary radius \emph{is} the Hill radius (we have previously assumed the Hill radius is 2.7 planetary radii).  This \edit1{true Roche Lobe Overflow (RLO)} scenario could manifest enormous mass loss rates, with the atmosphere's reservoir of mass fleeing the gravitational potential well directly, without the cushion of an exosphere.

\edit1{We propose that this terminal RLO should have a subtle-but-distinctive observational consequence.  Since heavy elements have smaller scale heights than hydrogen/helium, the observed ratio of heavy element to lighter element lines should be inversely correlated with the number of scale heights separating the white-light radius from the Hill radius.  All else equal, we would expect an anomalously large ratio of heavy metals compared to hydogen/Helium in the RLO case.  The extremely large Helium excess absorption, combined with the possibility of enhanced metal line depths in HAT-P-67~b---however uncertain---motivates a search for metal-line spectral tracers that may have previously been dismissed as too unlikely to detect.}
